---
title: "Project_1"
author: "Hana Kysela"
date: "May 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Project 1 - Simulation excercise

## Overview
In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem.

I am going to 
Show the sample mean and compare it to the theoretical mean of the distribution.
Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
Show that the distribution is approximately normal.

## Simulation


```{r}

set.seed(123456) #allows reproducibility

# packages management
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(ggplot2, knitr)

# The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 
# The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 
# Set lambda = 0.2 for all of the simulations. 
# You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

lambda <- 0.2   # lambda
n <- 40         # exponentials
sims <- 1000     # simulations

Exp <- matrix(data=rexp(n * sims, lambda), nrow=sims)
ExpMeans <- data.frame(means=apply(Exp, 1, mean))

# plot + more info
hist(ExpMeans$means)

  
```


#### STEAL FROM GITHUB ####

# https://github.com/TomLous/coursera-statistical-inference/blob/master/assessments/courseproject1.Rmd
lambda <- 0.2 # lambda for rexp
n <- 40 # number of exponetials
numberOfSimulations <- 1000 # number of tests

# set the seed to create reproducability
set.seed(11081979)

exponentialDistributions <- matrix(data=rexp(n * numberOfSimulations, lambda), nrow=numberOfSimulations)
exponentialDistributionMeans <- data.frame(means=apply(exponentialDistributions, 1, mean))

ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1) +   
  scale_x_continuous(breaks=round(seq(min(exponentialDistributionMeans$means), max(exponentialDistributionMeans$means), by=1)))


