---
title: "Project 1"
author: "Hana Kysela"
date: "May 21, 2016"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.height=4)
```


# Project 1 - Simulation excercise


## Overview

In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem.

I am going to 
Show the sample mean and compare it to the theoretical mean of the distribution.
Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
Show that the distribution is approximately normal.



## Simulation

```{r }

set.seed(1234) #allows reproducibility

# packages management
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(ggplot2, knitr)

# The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. 
# The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 
# Set lambda = 0.2 for all of the simulations. 
# You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

lambda <- 0.2   # lambda
n <- 40         # exponentials
sims <- 1000     # simulations

ExpMeans<-data.frame(sapply((1:sims), function(x) {mean(rexp(n, lambda))}))
colnames(ExpMeans)<- "means"

```



## Histogram of the means

```{r }

ggplot(data = ExpMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.2, fill = "grey") +  
  scale_x_continuous(breaks=round(seq(min(ExpMeans$means), max(ExpMeans$means), by=1)))
  
```

## 1. Show the sample mean and compare it to the theoretical mean of the distribution.

Expected (theoretical) mean $\mu$ is $\frac{1}{\lambda}$
```{r}
1/lambda
```
The actual mean from the simulation is 
```{r}
mean(ExpMeans$means)
```
## Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

Expected standard deviation of sample with standard deviation $\sigma$ is $\sigma = \frac{1/\lambda}{\sqrt{n}}$ 
Expected (theoretical) variance $Var$ of this sample is $Var = \sigma^2$

```{r}
sd<-1/lambda/sqrt(n)
var<-sd^2
var
```

The actual variance from the simulation is 
```{r}
sd<-sd(ExpMeans$means)
var<-sd^2
var

```
## 3. Show that the distribution is approximately normal.

The chart shows the histogram of simulated means and the theoretical values are in red, the simulation is in blue.

```{r }
ggplot(data = ExpMeans, aes(x = means)) + 
  geom_histogram(aes(y=..density..), binwidth = 0.2, fill = "grey") +
  stat_function(fun = dnorm, args = list(mean = 5 , sd = sd(ExpMeans$means)), colour = "red", size=1) + #theory
  stat_function(fun = dnorm, args = list(mean = mean(ExpMeans$means) , sd = sd(ExpMeans$means)), colour = "blue", size=1) + #sim
  geom_vline(xintercept = 5, size=1, colour="red") + # mean theory
  geom_vline(xintercept = mean(ExpMeans$means), size=1, colour="blue") + # mean sim

  scale_x_continuous(breaks=round(seq(min(ExpMeans$means), max(ExpMeans$means), by=1)))
  
```
